---
title: "a3_task3_travis_christy"
author: "Travis Christy"
date: "11/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(janitor)
library(here)
library(kableExtra)
library(lubridate)
```


**CITATION**: Reed, D. 2020. SBC LTER: Reef: Long-term experiment: Kelp removal: Urchin size frequency distribution ver 20. Environmental Data Initiative. [URL LINK](https://doi.org/10.6073/pasta/5a1e9ef03aa47bd2225c0bb98a02a63b) 


#### Task 3 Section 1: Size comparisons of red and purple urchins.

Reading in urchins.csv using here()

```{r}
urchins_tc <- read_csv(here("data", "urchins.csv")) %>% 
  clean_names()
```

##### Section 1.i Exploratory Data Visualization

Distribution of sizes between red and purple urchins represented by histograms and boxplots.

```{r}
ggplot(data = urchins_tc, aes(x = size)) +
  geom_histogram(bins = 20) +
  labs(x = "Urchin Diameter (cm)", y = "Total Urchins") +
  facet_wrap(~common_name)
```


```{r}
ggplot(data = urchins_tc, aes(x = common_name, y = size)) +
  geom_boxplot() +
  labs(x = "Species", y = "Urchin Diameter (cm)")
```


##### Section 1.ii Descriptive statistics table:

Using data wrangling techniques (group_by and summarize) to get a data set containing summary statistics for red and purple urchins.

```{r}
urchins_summary <- urchins_tc %>% 
  group_by(common_name) %>% 
  summarize(mean_size = mean(size, na.rm = TRUE),
            median_size = median(size, na.rm = TRUE),
            sd_size = sd(size, na.rm = TRUE),
            sample_size = n())

kbl(urchins_summary) %>% 
   kable_material("hover", full_width = F)

p_urchin_median_size <- urchins_summary[1,3]
r_urchin_median_size <- urchins_summary[2,3]

urchin_actual_size_diff <- r_urchin_median_size - p_urchin_median_size
urchin_median_size_diff <- 100*(r_urchin_median_size - p_urchin_median_size)/r_urchin_median_size
```

##### Section 1.iii Key takeways for size comparisons between the two urchin species:

- The histogram for purple urchins looks to have a combination distribution that is positively skewed. The kurtosis is leptokurtic, suggesting less dispersion in the purple urchin population size compared to the red urchin population.
- The histogram for red urchins shows a relatively normal distribution, however a flatter curve suggests the population is more dispersed in size than it is for purple urchins.
- The median size value (central tendency) for purple urchins vs red urchins (`r p_urchin_median_size` cm vs `r r_urchin_median_size` cm) suggests that on average red urchins are `r round(urchin_median_size_diff)`% larger or a difference of `r urchin_actual_size_diff` cm.
- According to the standard deviation values in the data tables, we can confirm what we saw in the graphs -- red urchin size is more dispersed than purple urchin size throughout the population.


#### Section 2. Explore purple urchin counts by site and year.

##### Section 2.i

```{r}
purple_urchins <- urchins_tc %>% 
  filter(common_name == "Purple Urchin") %>% 
  mutate(Date = mdy(date)) %>% 
  mutate(year = year(Date)) %>% 
  mutate(site_full = case_when(
    site == "NAPL" ~ "Naples Reef",
    site == "AQUE" ~ "Arroyo Quemado Reef",
    site == "CARP" ~ "Carpinteria Reef",
    site == "MOHK" ~ "Mohawk Reef",
    site == "IVEE" ~ "Isla Vista Reef"))
```

##### Section 2.ii

```{r}
purple_urchins_counts <- purple_urchins %>% 
  count(year, site_full)

kbl(purple_urchins_counts) %>% 
   kable_material("hover", full_width = F)
```

#### Section 2.iii

```{r}
ggplot(data = purple_urchins_counts, aes(x = year, y = n, color = site_full)) +
  geom_line() +
  theme_minimal()
```











